Improved Global Routing By Using A Star Algorithm

### 2020-06-01 paper.tex
Indentation

\pgfplotsset{
    % #1: index in the group(0,1,2,...)
    % #2: number of plots of that group
    bar group size/.style 2 args={
        /pgf/bar shift={%
                % total width = n*w + (n-1)*skip
                % -> subtract half for centering
                -0.5*(#2*\pgfplotbarwidth + (#2-1)*\pgfkeysvalueof{/pgfplots/bar group skip})  + 
                % the '0.5*w' is for centering
                (.5+#1)*\pgfplotbarwidth + #1*\pgfkeysvalueof{/pgfplots/bar group skip}},%
    },
    bar group skip/.initial=2pt,
    plot 0/.style={blue,fill=blue!30!white,mark=none},%
    plot 1/.style={red,fill=red!30!white,mark=none},%
    plot 2/.style={brown!60!black,fill=brown!30!white,mark=none},%
    plot 3/.style={brown!60!black,fill=brown!30!white,mark=none},%
}

@@

    \node [box,fill=green] at (0,0.5){A};
    \node[box,fill=green  ] at (2,2){A};  
    \node[box,fill=green  ] at (0,1){};  
    \node[box,fill=green  ] at (0,1.5){};  
    \node[box,fill=green  ] at (0,2){};  
    \node[box,fill=green  ] at (0.5,2){};  
    \node[box,fill=green  ] at (1,2){};  
    \node[box,fill=green  ] at (1.5,2){};  
    \node[box,fill=red  ] at (1.5,2.5){B}; 
    \node [box,fill=red] at (2.5,0.5){B}; 
    \node [box,fill=red] at (2.5,1){};
    \node [box,fill=red] at (2.5,1.5){};
    \node [box,fill=red] at (2.5,2){};
    \node [box,fill=red] at (2.5,2.5){};
    \node [box,fill=red] at (2,2.5){};
    \end{tikzpicture}
}
\caption{Finding path in global routing}
\label{fig:GlobalRoutingProcess}
\end{figure}

Figure~\ref{fig:GlobalRoutingProcess} shows a very simple approach of how the global router works. At (a) the source and target of both (A,B) need to be connected ignoring obstacles, nevertheless we can observe that the global router have to iterate to get the best routing paths, at (b) (B,B) connected but there is no way to connect (A,A) as when (B,B) connected together they blocked the way for (A,A) to be connected, after some iterations we can find the figure at (c) in which (A,A) and (B,B) connected correctly, but there is a problem, the (A,A) connection is not the optimal path as there are paths which achieve less wire length, so the global router have to iterate until reaching the optimal path. These iterations are done for only two connections in one layer without obstacles, then how about millions of wires in VLSI? this shows how much the global routing algorithm has to be very fast in order to connect this huge number of wires as fast as possible.

@@

@@

\section{Proposed Approach}
For global routing, three parameters are considered. The first is the total wire length, the second is the total overflow. The third is the running time. In our proposed solution, we are focused on the time parameter without affecting much the total wire length and total overflow. So according to minimize the time we have decided to apply A* algorithm as the pathfinder algorithm with a proposed heuristic function, A* is basically a guided variation of Dijkstra. This algorithm can be turned into a better or worse pathfinding algorithm by experimenting with the heuristics it uses and how it evaluates each node a* expands on a node only if it seems promising. Its only focus is to reach the goal node as quickly as possible from the current node, not to try and reach every other node. Illustrative example:
\begin{figure} [hbt!]
    \centering  
    \subfigure[Maze] {
    \begin{tikzpicture}
        [
            box/.style={rectangle,draw=black,thick, minimum size=0.5cm},
        ]

    \foreach \x in {0,0.5,...,3.5}{
        \foreach \y in {0,0.5,...,2}
            \node[box] at (\x,\y){};
    }
    \node [box,fill=green] at (0.5,1){S};
    \node[box,fill=green  ] at (3,1){D};  
    \node[box,fill=black  ] at (2,1){ };  
    \end{tikzpicture}
}

@@

    \node [box,fill=blue] at (0,0){};
    \node [box,fill=blue] at (0.5,0){};
    \node [box,fill=blue] at (1,0){};
    \node [box,fill=blue] at (1.5,0){};
    \node [box,fill=blue] at (2,0){};
    \node [box,fill=blue] at (2.5,0){};
    \node [box,fill=blue] at (3,0){};
    \node [box,fill=blue] at (0,0.5){};
    \node [box,fill=blue] at (0.5,0.5){};
    \node [box,fill=blue] at (1,0.5){};
    \node [box,fill=blue] at (1.5,0.5){};
    \node [box,fill=blue] at (2,0.5){};
    \node [box,fill=blue] at (2.5,0.5){};
    \node [box,fill=blue] at (3,0.5){};
    \node [box,fill=blue] at (3.5,.5){};
    \node [box,fill=blue] at (2.5,1){};
    \node [box,fill=blue] at (0,1){};
    \node [box,fill=blue] at (2.5,0.5){};
    \node [box,fill=blue] at (0,2){};
    \node [box,fill=blue] at (0.5,2){};
    \node [box,fill=blue] at (1,2){};
    \node [box,fill=blue] at (1.5,2){};
    \node [box,fill=blue] at (2,2){};
    \node [box,fill=blue] at (2.5,2){};
    \node [box,fill=blue] at (3,2){};
    \node [box,fill=blue] at (3.5,2){};
    \node [box,fill=blue] at (3.5,1.5){};
    \node [box,fill=blue] at (0,1.5){};
    \node [box,fill=blue] at (0.5,1.5){};
    \node [box,fill=blue] at (1,1.5){};
    \node [box,fill=yellow] at (0.5,1){S};
    \node [box,fill=yellow] at (1,1){};
    \node [box,fill=yellow] at (1.5,1){};
    \node [box,fill=yellow] at (1.5,1.5){};
    \node [box,fill=yellow] at (2,1.5){};
    \node [box,fill=yellow] at (2.5,1.5){};
    \node [box,fill=yellow] at (3,1.5){};
    \node[box,fill=yellow  ] at (3,1){D};  
    \node[box,fill=black  ] at (2,1){ };  
    \end{tikzpicture}
}
\subfigure[Applying A*] {
    \begin{tikzpicture}
        [
            box/.style={rectangle,draw=black,thick, minimum size=0.5cm},
        ]

    \foreach \x in {0,0.5,...,3.5}{
        \foreach \y in {0,0.5,...,2}
            \node[box] at (\x,\y){};
    }
    \node [box,fill=yellow] at (0.5,1){S};
    \node[box,fill=yellow  ] at (3,1){D};
    \node [box,fill=yellow] at (0.5,1.5){};  
    \node [box,fill=yellow] at (1,1.5){};  
    \node [box,fill=yellow] at (1.5,1.5){};
    \node [box,fill=yellow] at (2,1.5){};
    \node [box,fill=yellow] at (2.5,1.5){};
    \node [box,fill=yellow] at (3,1.5){};
    \node [box,fill=blue] at (0,1){};
    \node [box,fill=blue] at (1,1){};
    \node [box,fill=blue] at (1.5,1){};
    \node[box,fill=black  ] at (2,1){ };  
    \end{tikzpicture}
}
\caption{Comparison between Dijkstra \& A* }
\label{fig:GlobalRoutingProcess}
\end{figure}

In the illustrative example the visited nodes are marked in blue color and the path is in yellow color.
Both have path of six cells which in our case they have optimal wire length but totally different running time as Dijkstra visits much more nodes than A*
, in the case of many target nodes, Dijkstra might be better in total wire length but much slower than A*, with an appropriate heuristic function we can minimize the time and try to reach optimal wire length. The heuristic function is a way of informing the search about the path to a target that provides an intelligent way of guessing which neighbor of the node is going to lead to the target faster. Hence, our heuristic function composed of Manhattan distance, layer index, and some constants.

\begin{multline*}
    h = \ln ((\mid p1 \cdot x-p2 \cdot x\mid+\mid p1 \cdot y-p2 \cdot y\mid)*const1) \\ 
     +(\mid p1 \cdot layerid-p2 \cdot layerid\mid)*const2 \\
\end{multline*}
The global router is implemented using C++ with the boost library, and a parser for LEF/DEF formats. We conducted our experiment with the same configuration and tried to test the global router behavior using A* algorithm instead of Dijkstra's.

@@

\subsection{Proposed Algorithm}
% \floatstyle{norules}
\restylefloat{algorithm}
\begin{algorithm}[H]
    \caption{A* Algorithm}
    \hspace*{\algorithmicindent} \textbf{Input:} Grid values $g$, Source $s$, Destination $d$\\
    \hspace*{\algorithmicindent} \textbf{Output:} Path\\
    \begin{algorithmic}[1]
    \Procedure {A*}{}
    % Return  Average waiting time, throughput
    \Statex Declaration and Initialisation:
    \State Cells $\theta$ % The Cells in the grid.
    \State Visited Nodes $\phi$ % Nodes already visited.
    \State Openlist Nodes $\delta$ % Nodes still to be visited. pq
    \State The cost from starting point to a given node $g$
    \State The heuristic function $h$
    \State (The cost of $g$ + The cost of $h$) is $f$
    \If{$s$ or $d$ is invalid}
        \Return
    \EndIf
    \If{$s$ or $d$ is blocked}
        \Return
    \EndIf
    \If{$s$ = $d$}
        \Return
    \EndIf
    \State Initialize $\delta$ 
    \State Initialize $\phi$
        \While{$\delta$ is not empty}
            \State $currNode$  = node with least $f$ in $\delta$
            \State $\delta$.Remove($currNode$)
            \State Generate 4 neighbours of $currNode$
            \For{$neighbour \in neighbours$}
            \If{$neighbour$ = $d$}
                \State Get path
                \Return
            \ElsIf{$neighbour \notin \phi$ and $neighbour$ is not blocked}
                \State $gnew$ = $currNode$.$g$ + distance between neighbour and currNode
                \State $hnew$ = distance from $d$ to $neighbour$
                \State $fnew$ = $gnew$ + $hnew$
                \If{$fnew < neighbour$.$f$}
                \State $neighbour$.$g$ = $currNode$.$g$ + distance between neighbour and currNode
                \State $neighbour$.$h$ = distance from $d$ to $neighbour$
                \State $neighbour$.$f$ = $neighbour$.$g$ + $neighbour$.$h$
                \State $\delta$.Add($neighbour$)
                \EndIf
            \EndIf
            \EndFor
        \State $\phi$.Add($currNode$)
        \EndWhile
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

@@

    \addplot[plot 0,bar group size={0}{3}, red]
    coordinates {(Exp. 1,115) (Exp. 2,115) (Exp. 3,151) (Exp. 6,113.2) (Exp. 7,106.8)};
    \addplot[plot 1,bar group size={1}{3}, green]
    coordinates {(Exp. 1,150.8)};
    \addplot[plot 2,bar group size={2}{3}, blue]
    coordinates {(Exp. 1,161.4)};

@@

    \addplot[plot 0,bar group size={0}{3}, red]
    coordinates {(Exp. 1,68.8) (Exp. 2,68.8) (Exp. 3,88.7) (Exp. 6,62.4) (Exp. 7,63.4)};
    \addplot[plot 1,bar group size={1}{3}, green]
    coordinates {(Exp. 1,107.6)};
    \addplot[plot 2,bar group size={2}{3}, blue]
    coordinates {(Exp. 1,116.2)};

>>>>

\pgfplotsset{
    % #1: index in the group(0,1,2,...)
    % #2: number of plots of that group
    bar group size/.style 2 args={
        /pgf/bar shift={%
                % total width = n*w + (n-1)*skip
                % -> subtract half for centering
                -0.5*(#2*\pgfplotbarwidth + (#2-1)*\pgfkeysvalueof{/pgfplots/bar group skip})  + 
                % the '0.5*w' is for centering
                (.5+#1)*\pgfplotbarwidth + #1*\pgfkeysvalueof{/pgfplots/bar group skip}},%
    },
    bar group skip/.initial=2pt,
    plot 0/.style={cyan,fill=cyan!30!white,mark=none},%
    plot 1/.style={red,fill=red!30!white,mark=none},%
    plot 2/.style={brown!60!black,fill=brown!30!white,mark=none},%
    plot 3/.style={brown!60!black,fill=brown!30!white,mark=none},%
}

@@

    \node [box,fill=green] at (0,0.5){A};
    \node[box,fill=green  ] at (2,2){A};  
    \node[box,fill=green  ] at (0,1){};  
    \node[box,fill=green  ] at (0,1.5){};  
    \node[box,fill=green  ] at (0,2){};  
    \node[box,fill=green  ] at (0.5,2){};  
    \node[box,fill=green  ] at (1,2){};  
    \node[box,fill=green  ] at (1.5,2){};  
    \node[box,fill=red  ] at (1.5,2.5){B}; 
    \node [box,fill=red] at (2.5,0.5){B}; 
    \node [box,fill=red] at (2.5,1){};
    \node [box,fill=red] at (2.5,1.5){};
    \node [box,fill=red] at (2.5,2){};
    \node [box,fill=red] at (2.5,2.5){};
    \node [box,fill=red] at (2,2.5){};
    \end{tikzpicture}
}
\caption{Finding path in global routing}
\label{fig:GlobalRoutingProcess}
\end{figure}
\medskip
Figure~\ref{fig:GlobalRoutingProcess} shows a very simple approach of how the global router works. At (a) the source and target of both (A,B) need to be connected ignoring obstacles, nevertheless we can observe that the global router have to iterate to get the best routing paths, at (b) (B,B) connected but there is no way to connect (A,A) as when (B,B) connected together they blocked the way for (A,A) to be connected, after some iterations we can find the figure at (c) in which (A,A) and (B,B) connected correctly, but there is a problem, the (A,A) connection is not the optimal path as there are paths which achieve less wire length, so the global router have to iterate until reaching the optimal path. These iterations are done for only two connections in one layer without obstacles, then how about millions of wires in VLSI? this shows how much the global routing algorithm has to be very fast in order to connect this huge number of wires as fast as possible.

@@

\medskip

@@

\section{Proposed Approach}
For global routing, three parameters are considered. The first is the total wire length, the second is the total overflow. The third is the running time. In our proposed solution, we are focused on the time parameter without affecting much the total wire length and total overflow. So according to minimize the time we have decided to apply A-Star algorithm as the pathfinder algorithm with a proposed heuristic function, A-Star is basically a guided variation of Dijkstra. This algorithm can be turned into a better or worse pathfinding algorithm by experimenting with the heuristics it uses and how it evaluates each node A-Star expands on a node only if it seems promising. Its only focus is to reach the goal node as quickly as possible from the current node, not to try and reach every other node. Illustrative example:
\begin{figure} [hbt!]
    \centering  
    \subfigure[Maze] {
    \begin{tikzpicture}
        [
            box/.style={rectangle,draw=black,thick, minimum size=0.5cm},
        ]

    \foreach \x in {0,0.5,...,3.5}{
        \foreach \y in {0,0.5,...,2}
            \node[box] at (\x,\y){};
    }
    \node [box,fill=orange] at (0.5,1){S};
    \node[box,fill=orange  ] at (3,1){D};  
    \node[box,fill=black  ] at (2,1){ };  
    \end{tikzpicture}
}

@@

    \node [box,fill=cyan] at (0,0){};
    \node [box,fill=cyan] at (0.5,0){};
    \node [box,fill=cyan] at (1,0){};
    \node [box,fill=cyan] at (1.5,0){};
    \node [box,fill=cyan] at (2,0){};
    \node [box,fill=cyan] at (2.5,0){};
    \node [box,fill=cyan] at (3,0){};
    \node [box,fill=cyan] at (0,0.5){};
    \node [box,fill=cyan] at (0.5,0.5){};
    \node [box,fill=cyan] at (1,0.5){};
    \node [box,fill=cyan] at (1.5,0.5){};
    \node [box,fill=cyan] at (2,0.5){};
    \node [box,fill=cyan] at (2.5,0.5){};
    \node [box,fill=cyan] at (3,0.5){};
    \node [box,fill=cyan] at (3.5,.5){};
    \node [box,fill=cyan] at (2.5,1){};
    \node [box,fill=cyan] at (0,1){};
    \node [box,fill=cyan] at (2.5,0.5){};
    \node [box,fill=cyan] at (0,2){};
    \node [box,fill=cyan] at (0.5,2){};
    \node [box,fill=cyan] at (1,2){};
    \node [box,fill=cyan] at (1.5,2){};
    \node [box,fill=cyan] at (2,2){};
    \node [box,fill=cyan] at (2.5,2){};
    \node [box,fill=cyan] at (3,2){};
    \node [box,fill=cyan] at (3.5,2){};
    \node [box,fill=cyan] at (3.5,1.5){};
    \node [box,fill=cyan] at (0,1.5){};
    \node [box,fill=cyan] at (0.5,1.5){};
    \node [box,fill=cyan] at (1,1.5){};
    \node [box,fill=orange] at (0.5,1){S};
    \node [box,fill=orange] at (1,1){};
    \node [box,fill=orange] at (1.5,1){};
    \node [box,fill=orange] at (1.5,1.5){};
    \node [box,fill=orange] at (2,1.5){};
    \node [box,fill=orange] at (2.5,1.5){};
    \node [box,fill=orange] at (3,1.5){};
    \node[box,fill=orange  ] at (3,1){D};  
    \node[box,fill=black  ] at (2,1){ };  
    \end{tikzpicture}
}
\subfigure[Applying A-Star] {
    \begin{tikzpicture}
        [
            box/.style={rectangle,draw=black,thick, minimum size=0.5cm},
        ]

    \foreach \x in {0,0.5,...,3.5}{
        \foreach \y in {0,0.5,...,2}
            \node[box] at (\x,\y){};
    }
    \node [box,fill=orange] at (0.5,1){S};
    \node[box,fill=orange  ] at (3,1){D};
    \node [box,fill=orange] at (0.5,1.5){};  
    \node [box,fill=orange] at (1,1.5){};  
    \node [box,fill=orange] at (1.5,1.5){};
    \node [box,fill=orange] at (2,1.5){};
    \node [box,fill=orange] at (2.5,1.5){};
    \node [box,fill=orange] at (3,1.5){};
    \node [box,fill=cyan] at (0,1){};
    \node [box,fill=cyan] at (1,1){};
    \node [box,fill=cyan] at (1.5,1){};
    \node[box,fill=black  ] at (2,1){ };  
    \end{tikzpicture}
}
\caption{Comparison between Dijkstra \& A-Star }
\label{fig:GlobalRoutingProcess}
\end{figure}
\medskip
In the illustrative example the visited nodes are marked in cyan color, the path is in orange color, and the blocked cells are black.
Both have path of six cells which in our case they have optimal wire length but totally different running time as Dijkstra visits much more nodes than A-Star
, in the case of many target nodes, Dijkstra might be better in total wire length but much slower than A-Star, with an appropriate heuristic function we can minimize the time and try to reach optimal wire length. The heuristic function is a way of informing the search about the path to a target that provides an intelligent way of guessing which neighbor of the node is going to lead to the target faster. Hence, our heuristic function composed of Manhattan distance, layer index, and some constants.

\begin{multline*}
    h = \ln ((\mid p1 \cdot x-p2 \cdot x\mid+\mid p1 \cdot y-p2 \cdot y\mid)*const1) \\ 
     +(\mid p1 \cdot layerid-p2 \cdot layerid\mid)*const2 \\
\end{multline*}
The global router is implemented using C++ with the boost library, and a parser for LEF/DEF formats. We conducted our experiment with the same configuration and tried to test the global router behavior using A-Star algorithm instead of Dijkstra's.

@@

\subsection{Proposed Algorithm}
% \floatstyle{norules}
\restylefloat{algorithm}
\begin{algorithm}[H]
    \caption{A-Star Algorithm}
    \hspace*{\algorithmicindent} \textbf{Input:} Grid values $g$, Source $s$, Destination $d$\\
    \hspace*{\algorithmicindent} \textbf{Output:} Path\\
    \begin{algorithmic}[1]
    \Procedure {A-Star}{}
    % Return  Average waiting time, throughput
    \Statex Declaration and Initialisation:
    \State Cells $\theta$ % The Cells in the grid.
    \State Visited Nodes $\phi$ % Nodes already visited.
    \State Openlist Nodes $\delta$ % Nodes still to be visited. pq
    \State The cost from starting point to a given node $g$
    \State The heuristic function $h$
    \State (The cost of $g$ + The cost of $h$) is $f$
    \If{$s$ or $d$ is invalid}
        \Return
    \EndIf
    \If{$s$ or $d$ is blocked}
        \Return
    \EndIf
    \If{$s$ = $d$}
        \Return
    \EndIf
    \State Initialize $\delta$ 
    \State Initialize $\phi$
        \While{$\delta$ is not empty}
            \State $currNode$  = node with least $f$ in $\delta$
            \State $\delta$.Remove($currNode$)
            \State Generate 4 neighbours of $currNode$
            \For{$neighbour \in neighbours$}
            \If{$neighbour$ = $d$}
                \State Get path
                \Return
            \ElsIf{$neighbour \notin \phi$ and $neighbour$ is not blocked}
                \State $gnew$ = $currNode$.$g$ + distance between neighbour and currNode
                \State $hnew$ = distance from $d$ to $neighbour$
                \State $fnew$ = $gnew$ + $hnew$
                \If{$fnew < neighbour$.$f$}
                \State $neighbour$.$g$ = $currNode$.$g$ + distance between neighbour and currNode
                \State $neighbour$.$h$ = distance from $d$ to $neighbour$
                \State $neighbour$.$f$ = $neighbour$.$g$ + $neighbour$.$h$
                \State $\delta$.Add($neighbour$)
                \EndIf
            \EndIf
            \EndFor
        \State $\phi$.Add($currNode$)
        \EndWhile
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

@@

    \addplot[plot 0,bar group size={0}{3}, red]
    coordinates {(Exp. 1,115) (Exp. 2,115) (Exp. 3,151) (Exp. 6,113.2) (Exp. 7,106.8)};
    \addplot[plot 1,bar group size={1}{3}, green]
    coordinates {(Exp. 1,150.8)};
    \addplot[plot 2,bar group size={2}{3}, cyan]
    coordinates {(Exp. 1,161.4)};

@@

    \addplot[plot 0,bar group size={0}{3}, red]
    coordinates {(Exp. 1,68.8) (Exp. 2,68.8) (Exp. 3,88.7) (Exp. 6,62.4) (Exp. 7,63.4)};
    \addplot[plot 1,bar group size={1}{3}, green]
    coordinates {(Exp. 1,107.6)};
    \addplot[plot 2,bar group size={2}{3}, cyan]
    coordinates {(Exp. 1,116.2)};
